---
title: "Computation for Analytics 1"
output: html_notebook
runtime: shiny
---
By:  Xavier M. Puspus  
Introduction to R: Homework 2

+ Upload Assignment01.rmd to github.
+ Define an R function that removes NA values from a vector.
+ Define an R function that computes the factorial of given an integer argument. The
output should be a vector of length 1.
+ Define an R function that computes the determinant of a given matrix. The output
should be a vector of length 1. (use basic r and operations)
+ Define an R function that sorts a given vector in decreasing order. The output should be
a vector of the same length. It should accept both numeric or character vectors.
+ Define an R function that accepts a POSIXct as argument and outputs the day of the
week as characters. (Don't use special functions. Instead use modulo and define weekdays)
+ Create a function to compute for your net pay at work (input: # of days worked, tax classification, output: 1 number)
+ Create a function to return the nth highest number


### i. Upload Assignment01.rmd to Github.
Please see [repo](https://github.com/xmpuspus/RCode/blob/master/Assignment01.Rmd) for this submission.

### 1. Removes NA values from a vector. 
```{r}

## Easy Way
# remove_na = function(d){d[!is.na(d)]}

## Harder way (with loops)
rem_na <- function(sample_vec) {
  init <- c()
  for (i in 1:length(sample_vec)) {
    if (!is.na(sample_vec[1])){
      init <- c(init, sample_vec[i])
    }
  }
  return (init)
}


```
### 2. Computes the factorial of given an integer argument.  
The output should be a vector of length 1. I used recursive calculation of *n* since its fast but memory intensive.

```{r}
## Answer to Homework is the function below
factorial = function(x) {
  # if x in [0,1], return 1
  if (x == 0 || x == 1) {
return (1)
} 
  else {
return (x * factorial(x - 1)) 
} 
}
```

### 3. Sorts a given vector in decreasing order.  
Should accept numeric and character vectors.
```{r}

# arr = c(1,2,10,7,9, 100, 1000, -1, 3,4) # numeric
arr = c('ab','bc','dd','wz', 'ef') # characters

# insertion sort algorithm 
ins_sort <- function(arr){
  for (j in 2:length(arr)) {
    init = arr[j] 

    i = j - 1 
    while (i > 0 && arr[i] < init) {
      arr[(i + 1)] = arr[i]
      i = i - 1 
    }
    arr[(i + 1)] = init
  }
  return (arr)
} 

# ins_sort(arr)

```


### 4. Return the nth highest number  
I used the insertion sort function from 4.
```{r}

# get nth index from result of insertion_sort  
nth_largest <- function(arr, nth){
  return (ins_sort(arr)[nth])
} 

# nth_largest(arr, 3)

```

### 5. Create a function to compute for your net pay at work.  
We compute the monthly net pay based on:    

+ monthly basic pay  
+ taxable allowances  
+ non-taxable allowances  
+ Leave without pay days  
+ months of payment (13 for PH)  
+ and number of working days (default is 22)  

```{r}
net_pay = function(basic, tax_alw = 0, ntax_alw = 0, lwop_days = 0, months_pay = 13, working_days = 22){
  
  annual = (basic + tax_alw) * months_pay

  if (annual <=  250000){
    net = annual
  } else if (annual <= 400000) {
    net = annual - (annual - 250000) * 0.2
  } else if (annual <= 800000) {
    net = annual - (annual - 400000) * 0.25 - 30000
  } else if (annual <= 2000000) {
    net = annual - (annual - 800000) * 0.30 - 130000
  } else if (annual <= 8000000) {
    net = annual - (annual - 2000000) * 0.32 - 490000
  } else {
    net = annual - (annual - 8000000) * 0.35 - 2410000
  }
    
  mnet = net * 1./months_pay
  net_final = mnet + ntax_alw - (mnet * 1./working_days) * lwop_days
  return (net_final)
}

net_pay(50000, 1000, 1000, 2, 13, 20)
```

# Shiny Applications
This is an interactive application that allows reactive adjustments of values where we instantly see the results. I built 2 apps, 1 for factorial and 1 for computing net pay.

### Factorials  
```{r echo = FALSE}
library(shiny)


ui <- fluidPage(
  headerPanel("Compute Factorial"),
  sidebarLayout(
  sidebarPanel(
  numericInput("n", "Number:", 10, min = 1, max = 100)
),
mainPanel(
  h4('Factorial is:'),
  verbatimTextOutput("value"))
 )
)

server <- function(input, output) {
  

  
  output$value <- renderText({
  print(factorial(input$n))

  })
}

shinyApp(ui = ui, server = server)

```
### Net Pay Calculator
```{r echo = FALSE}
library(shiny)


ui2 <- fluidPage(
  sidebarLayout(
  sidebarPanel(
  numericInput("basic", "Basic Pay:", 50000, step = 1000, min = 1000, max = 10000000),
  numericInput("tax_alw", "Taxable Allowance:", 0, step = 1000, min = 500, max = 10000000),
  numericInput("ntax_alw", "Non-taxable Allowance:", 0, step = 1000, min = 500, max = 10000000),
  numericInput("lwop_days", "Leave Without Pay Days:", 0, step = 1, min = 0, max = 31),
  numericInput("months_pay", "Months:", 13, step = 1, min = 1, max = 20),
  numericInput("working_days", "Working Days:", 22, step = 1, min = 1, max = 31)
),
mainPanel(
  h4('Net Monthly Salary is:'),
  verbatimTextOutput("value"))
 )
)

server2 <- function(input, output) {
  

  
  output$value <- renderText({
  print(net_pay(input$basic,input$tax_alw, input$ntax_alw, input$lwop_days, input$months_pay, input$working_days))

  })
}

shinyApp(ui = ui2, server = server2)

```




