---
title: "Computation for Analytics 1"
output: html_notebook
---
By:  Xavier M. Puspus  
Intro to R: Homework 1

This markdown document shows the code as well as the output for Homework 1 on Introduction to R.

# 1. WHO Dataset

Find out:  
  + country with the biggest population,
  + population of Malaysia,
  + Country with the lowest literacy,  
  + Richest country in Europe based on GNI,  
  + Mean life expetancy of countries in Africa,  
  + Number of countries with population greater than 10,000,  
  + Top 5 countries in the Americas with the highest child mortality.  
  

### Read Data  
Load WHO dataset and display first 6 rows.
```{r echo = TRUE}

# Load Data
who_df = read.csv('data/WHO.csv')
head(who_df)
```
### Data of Country with the biggest population  
Metadata of country with biggest population.
```{r echo = TRUE}
who_df[which.max(who_df$Population),]
```
### Country with the biggest population
```{r}
who_df[which.max(who_df$Population),]$Country
```
### Country with Lowest Literacy Rate  
```{r echo = TRUE}
# Get columns we need
who_df_lit = who_df[c('Country', 'LiteracyRate')]

# Remove NAs
who_df_lit_nona = who_df_lit[complete.cases(who_df_lit),]

# Get bottom 10
bottom_10_lit = tail(who_df_lit_nona[order(who_df_lit_nona$LiteracyRate, decreasing = TRUE),], 10)

# Get Country with lowest literacy rate
bottom_10_lit[10,][1]
```
Another method of getting country with lowest literacy rate:
```{r}
who_df[which.min(who_df$LiteracyRate),]$Country
```


### Richest Country in Europe Based on GNI
```{r}
# Get columns we need filtering for Region = Europe
who_df_gni = who_df[which(who_df$Region == "Europe"),][c('Country', 'GNI')]

# Remove NAs
who_df_gni_nona = who_df_gni[complete.cases(who_df_gni),]

# Get top 10
top_10_gni = head(who_df_gni_nona[order(who_df_gni_nona$GNI, decreasing = TRUE),], 10)

#Get Country with lowest literacy rate

top_10_gni[1,][1]
```
Another Method of getting richest country in Europe based on GNI:
```{r}
who_df[which(who_df$Region == "Europe"),][which.max(who_df[which(who_df$Region == "Europe"),]$GNI),]$Country
```

### Mean life expetancy of countries in Africa
```{r echo = TRUE}
# filter for the columns that we need and for region = Africa
who_df_le = who_df[which(who_df$Region == "Africa"),][c('Country', 'LifeExpectancy')]

# Get mean of life expectancy
mean(who_df_le$LifeExpectancy, na.rm = TRUE)
```
### Countries with population greater than 10,000
```{r echo = TRUE}
who_df[which(who_df$Population > 10000),]
```
### Number of countries with population greater than 10,000
```{r echo = TRUE}
dim(who_df[which(who_df$Population > 10000),])[1]
```

### Data on Top 5 countries in the Americas with the highest child mortality
```{r echo = TRUE}
who_df_cm = who_df[which(who_df$Region == "Americas"),][c('Country', 'ChildMortality')]
who_df_cm_nona = who_df_cm[complete.cases(who_df_cm),]
top_10_cm = head(who_df_cm_nona[order(who_df_cm_nona$ChildMortality, decreasing = TRUE),], 5)
top_10_cm
```
### Top 5 countries in the Americas with the highest child mortality
```{r}
top_10_cm$Country
```

# 2. Historical NBA Performance

Find out:  
+ The year Bulls has the highest winning percentage,  
+ Teams with an even win-loss record in a year.   


### Read Data
```{r echo = TRUE}
# Load Library
library("readxl")

# Read Data
data = read_excel('data/Historical NBA Performance.xlsx')

# Rename last column
names(data)[4]<-"win_pct"
head(data)
```

### The year the Bulls had the Highest Winning Percentage
```{r echo = TRUE}
# Who are the teams with the highest winning percentage each year? (using highest win percentage of all time)
bulls_data = data[which(data$Team == 'Bulls'),]
bulls_data[which.max(bulls_data$win_pct),]['Year']
```

### Data Augmented with Win-Loss Percentage Columns
Equal win-loss

```{r}
# Even Win-loss 

# Create table that splits win loss
wl = do.call(cbind, c(data, read.table(text = data$Record, sep = "-", colClasses = "character")))

# Convert to DataFrame
wl = as.data.frame.matrix(wl) 

# Convert last 2 columns to numeric
wl$V1 = as.numeric(as.character(wl$V1))
wl$V2 = as.numeric(as.character(wl$V2))

# Replace column names
names(wl)[5] = 'Win'
names(wl)[6] = 'Loss'

# Inspect data we've cleaned
head(wl)
```
### Data of teams with even win-loss
```{r echo = TRUE}
equal_wl = wl[which(wl$Win == wl$Loss),]
equal_wl[c("Year", "Team", "Win", "Loss")]
```
### Teams with an even win-loss
```{r echo = TRUE}
unique(equal_wl$Team)
```

# 3. Season Stats  

Find out:  
  + Player with the highest 3-pt attempt rate in a season.   
  + Player with the highest free throw rate in a season.   
  + What year/season does Lebron James scored the highest?   
  + What year/season does Michael Jordan scored the highest?   
  + Player efficiency rating of Kobe Bryant in the year where his MP is the lowest?   
  
### Read Data
```{r echo = TRUE}
data = read.csv('data/seasons_stats.csv')
data$Year = as.numeric(data$Year)
head(data)
```
### Data of Players with the Highest 3-Point Attempt Rate In a Season per Year
```{r echo = TRUE}
# Players with highest 3-point attempt rate in a season per year?
df_agg_X3PAr = do.call(rbind, lapply(split(data, as.factor(data$Year)), function(x) {return(x[which.max(x$X3PAr),])}))
df_agg_X3PAr[order(df_agg_X3PAr$Year, decreasing = FALSE),][c('Year', 'Player', 'X3PAr')]
```
### Data of Players with the Highest 3-Point Attempt Rate
```{r echo = TRUE}
df_agg_X3PAr[c("Player", 'X3PAr')][which(df_agg_X3PAr$X3PAr == max(data$X3PAr, na.rm = TRUE)),]
```
### Players with the Highest 3-Point Attempt Rate
```{r echo = TRUE}
unique(df_agg_X3PAr[c("Player", 'X3PAr')][which(df_agg_X3PAr$X3PAr == max(data$X3PAr, na.rm = TRUE)),]$Player)
```



### Data of Players with the Highest 3-Point Attempt Rate In a Season of All the Player's Time
```{r echo = TRUE}
data[c("Player", 'X3PAr')][which(data$X3PAr == max(data$X3PAr, na.rm = TRUE)),]
```
### Players with the Highest 3-Point Attempt Rate In a Season of All the Player's Time
```{r echo = TRUE}
unique(data[c("Player", 'X3PAr')][which(data$X3PAr == max(data$X3PAr, na.rm = TRUE)),]$Player)
```

### Data of Players with the Highest Free Throw Rate In a Season per Year
```{r echo = TRUE}
# Players with highest 3-point attempt rate in a season per year?
df_agg_FTr = do.call(rbind, lapply(split(data, as.factor(data$Year)), function(x) {return(x[which.max(x$FTr),])}))
df_agg_FTr[order(df_agg_FTr$Year, decreasing = FALSE),][c('Year', 'Player', 'FTr')]
```
### Data of Players with the Highest Free Throw Rate In a Season
```{r echo = TRUE}
df_agg_FTr[c("Player", 'FTr')][which(df_agg_FTr$FTr == max(df_agg_FTr$FTr, na.rm = TRUE)),]
```
### Players with the Highest Free Throw Rate In a Season
```{r}
df_agg_FTr[c("Player", 'FTr')][which(df_agg_FTr$FTr == max(df_agg_FTr$FTr, na.rm = TRUE)),]$Player
```

### Data of Players with the Highest Free Throw Rate in a Season All Player's Time
```{r echo = TRUE}
data[c("Player", 'FTr')][which(data$FTr == max(data$FTr, na.rm = TRUE)),]
```

### Players with the Highest Free Throw Rate in a Season All Player's Time
```{r echo = TRUE}
data[c("Player", 'FTr')][which(data$FTr == max(data$FTr, na.rm = TRUE)),]$Player
```

### Data of Players Who Scored the Highest Each Year
```{r echo = TRUE}
# Who are the players with the highest Scores each year?
df_agg_teams = do.call(rbind, lapply(split(data, as.factor(data$Year)), function(x) {return(x[which.max(x$PTS),])}))
df_agg_teams[order(df_agg_teams$Year, decreasing = FALSE),][c('Year', 'Player', 'PTS')]
```
### Years LeBron James Scored the Highest that Season
```{r echo = TRUE}
df_agg_teams[which(df_agg_teams$Player == 'LeBron James'),]
```
```{r echo = TRUE}
df_agg_teams[which(df_agg_teams$Player == 'LeBron James'),]$Year
```
Using this method of aggregation, we see that LeBron James has never had the highest score compared to other players for each year.
  
### Year LeBron James Scored his Highest Points of All Time  
For this type of aggregation, we get LeBron's max highest point and output the year he got this highest point.

```{r echo = TRUE}
data[which(data$Player == 'LeBron James'),][which.max(data[which(data$Player == 'LeBron James'),]$PTS),]
```
```{r echo = TRUE}
data[which(data$Player == 'LeBron James'),][which.max(data[which(data$Player == 'LeBron James'),]$PTS),]$Year
```

### Years Michael Jordan Scored the Highest that Season
```{r echo = TRUE}
df_agg_teams[which(df_agg_teams$Player == 'Michael Jordan*'),]
```
```{r echo = TRUE}
df_agg_teams[which(df_agg_teams$Player == 'Michael Jordan*'),]$Year
```
### Year Michael Jordan Scored His Highest Points of All Time
```{r echo = TRUE}
data[which(data$Player == 'Michael Jordan*'),][which.max(data[which(data$Player == 'Michael Jordan*'),]$PTS),]
```
```{r echo = TRUE}
data[which(data$Player == 'Michael Jordan*'),][which.max(data[which(data$Player == 'Michael Jordan*'),]$PTS),]$Year
```

### Player efficiency rating of Kobe Bryant in the year where his MP is the lowest?
```{r echo = TRUE}
kb_data = data[which(data$Player == 'Kobe Bryant'),]
kb_data[which.min(kb_data$MP),]$PER
```
# 4. National Universities Rankings  
Find Out:  
+ University with the most number of undergrads  
+ Average Tuition in the Top 10 University

### Read Data  
```{r echo = TRUE}
data = read.csv('data/National Universities Rankings.csv')
head(data)
```

### University with the most number of undergrads
```{r echo = TRUE}
data[which.max(data[['Undergrad.Enrollment']]),]$Name
```
### Top 10 Universities

Top 10 universities based on rank.
```{r echo = TRUE}
head(data[order(data[['Rank']], decreasing = FALSE),], 10)
```


### Average Tuition of the Top 10 Universities
```{r echo = TRUE}
# Get average tuition rates of to 10 univerisities
mean(as.numeric(gsub(",", "", sub('.', '', head(data[order(data[['Rank']], decreasing = FALSE),], 10)[['Tuition.and.fees']]))))
```
# END OF CODE
